---
title: "Distribution of job listings across Australia"
author:
- familyname: Troy
  othernames: Helen
  title: Professor
  qualifications: BSc, PhD
  email: helen.troy@gmail.com
  correspondingauthor: true
- familyname: Khan
  othernames: Genghis
  title: Mr
  qualifications: BA, BComm, BCompSci, BLaw
department: Department of\newline Econometrics &\newline Business Statistics
organization: Acme Corporation
bibliography: references.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  MonashEBSTemplates::report:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      cache=TRUE, 
                      messages=FALSE, 
                      warning=FALSE)
```

```{r}
library(ggplot2)
library(tidyverse)
library(haven)
library(lubridate)
library(broom)
library(tidyr)
library(base)
library(glue)
library(kableExtra)
library(bookdown)
```


# Introduction 

Job types, job locations, rising demand for certain job categories, these are just some of the aspects of job hunting that many working people around the world are intrigued by. In this report, you will find an in depth analysis into the distribution of job listings across Australia. The report will analyze 3 main sections of job distribution within Australia, broken down by:

1) The change in demand by job category
2) The change in demand for jobs by city
3) The change in demand by job type

Our motivation to choose this as our research topic is prompted by experiencing first hand the cataclysmic effects of the coronavirus and the impact it has had on the job market globally. It has been an unfortunate scenario to witness unemployment rates skyrocket in major cities around the world. Thus, we want to visualize the changes in demand to jobs of all types and categories in Australia, particularly in the major cities.

# Dataset and Methodology
- Source of data. (Dewi)  
- Limitations (Hanh - master)  
- Methodology: How we deal with this dataset. ( Peimin - master)  

# Analysis:
### Clean the data: 
1. isolate the month + year (Hanh - master branch)  
```{r load-data, cache = TRUE, echo=FALSE}
raw_data <- read.csv(here::here("data", "seek_australia_sample.csv"))
```

```{r clean-data, cache = TRUE, echo=FALSE}
# choose the needed columns only
data_cleaned <- raw_data %>%
  select(3:9)

# separate the post-date column to get the date out
data_cleaned <- data_cleaned %>%
  separate (post_date, into = c("post_date", "tz"), sep = "T")

# convert the post_date column from character to date format
data_cleaned %>%
  mutate_at(vars(post_date), as.Date, format = "%Y-%m-%d")
```

2. Missing values  (Hanh - Hanh-Ngo branch)  

### Analyze the change in demand by job category) (Peimin-Peimin branch)


### Analyze the change in demand by city + life quality  
1. Change in demand (Hanh)    
2. How does it reflect life quality? (Dewi)

### Analyze the change in demand by job type

In this section, our goal is to see the different types of jobs such as full-time, part-time, contract and casual jobs. The changes in demand to these job types will vastly differ across the months, but it will be interesting to see which job type increased in demand.

#### Number of jobs posted per job type

```{r job-type-counts, echo=FALSE}
data_cleaned %>% group_by(job_type) %>% summarise(`Number of jobs posted` = n()) %>% rename(`Job Type` = job_type) %>%
  kable(caption = "Number of jobs posted by job type") %>%
  kable_styling(bootstrap_options = c("hovered", "striped"))
```

```{r job-types, fig.align="center", fig.cap="Number of jobs for each job type}
data_cleaned %>% group_by(job_type) %>% count() %>% 
    ggplot(aes(x = n, y = job_type, fill = job_type)) +
      geom_col() +
      theme(legend.position = "none") +
      labs(x = "Number of jobs posted", y = "Job Type")
```

The different types of jobs being posted on seek are part time, full time, contract/temporary and casual/vacation. From Table \@ref(tab:job-type-counts), we can see that over 13500 jobs out of the roughly 20000 jobs posted are full-time jobs. It is easier to visualize just how common these types of jobs are in Figure \@ref(fig:job-types), and we can see how dominant the full-time jobs are being posted by companies. According to abc.net, most of the full-time jobs are being generated in the healthcare and education sectors. Specifically, the National Disability Insurance Scheme (NDIS) and aged care services have increased demand for these types of jobs. The NDIS expects an extra 80,000 full-time jobs by 2020.

```{r, echo=FALSE}
job_types <- data_cleaned %>% select(job_title, post_date, job_type) %>% 
                                     mutate(post_date = as.Date(post_date))
```

```{r, echo=FALSE}
job_type_counts <- job_types %>% mutate(Month = month(post_date),
                                       job_type = as.character(job_type)) %>%
                            group_by(Month, job_type) %>% summarise("Number of jobs posted" = n())

```

```{r, eval=FALSE}
data_cleaned %>% mutate(post_date = as.Date(post_date)) %>% group_by(post_date, job_type) %>% summarise("Number of jobs posted" = n()) %>% ungroup() %>% 
    ggplot(aes(x = post_date, y = `Number of jobs posted`)) +
      geom_line(aes(color = job_type)) +
      theme_classic() +
      labs(x = "Number of jobs posted", y = "Job Type")
```

#### Fluctuations in job posting numbers by month

```{r job-count-fluctations, fig.align="center", fig.cap="Changes in job number postings by month"}
ggplot(job_type_counts, aes(x = Month, y = `Number of jobs posted`)) +
  geom_point(aes(color = job_type)) +
  geom_line(aes(color= job_type)) 
```

However, the fluctuations of these job types says more about what kinds of jobs companies are willing to offer. For example, in Figure \#ref(fig:job-count-fluctations), there is a large spike in full time jobs between August and October. The opposite is true when these jobs plummet down from 3100 posted jobs in October to around 700 jobs in November. Both part time and casual jobs stay relatively neutral and do not change as much throughout the months. According to Aigroup, casual jobs have stayed the same since the past two decades, these jobs formed 20.6% of the workforce in February 2018 compared to 20.1 % in 1998. The reason for an uptick in the jobs posted for this category in October is because the largest proportion of people who take up casual jobs are aged 20-24. Thus around October, many of these people who are students are available as the summer break approaches. Interestingly, all the job types follow a similar pattern, reporting an increase in the number of jobs posted in June and October as these are the time periods for winter and summer break respectively.



# Conclusion: (Justin - master branch)




