---
title: "Distribution of job listings across Australia"
author:
- familyname: Troy
  othernames: Helen
  title: Professor
  qualifications: BSc, PhD
  email: helen.troy@gmail.com
  correspondingauthor: true
- familyname: Khan
  othernames: Genghis
  title: Mr
  qualifications: BA, BComm, BCompSci, BLaw
department: Department of\newline Econometrics &\newline Business Statistics
organization: Acme Corporation
bibliography: references.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  MonashEBSTemplates::report:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      cache=TRUE, 
                      messages=FALSE, 
                      warning=FALSE)
```

```{r}
library(ggplot2)
library(tidyverse)
library(haven)
library(lubridate)
library(broom)
library(tidyr)
library(base)
library(glue)
```


# Introduction 

Job types, job locations, rising demand for certain job categories, these are just some of the aspects of job hunting that many working people around the world are intrigued by. In this report, you will find an in depth analysis into the distribution of job listings across Australia. The report will analyze 3 main sections of job distribution within Australia, broken down by:

1) The change in demand by job category
2) The change in demand for jobs by city
3) The change in demand by job type

Our motivation to choose this as our research topic is prompted by experiencing first hand the cataclysmic effects of the coronavirus and the impact it has had on the job market globally. It has been an unfortunate scenario to witness unemployment rates skyrocket in major cities around the world. Thus, we want to visualize the changes in demand to jobs of all types and categories in Australia, particularly in the major cities.

# Dataset and Methodology
- Source of data. (Dewi)  
- Limitations (Hanh - master)  
- Methodology: How we deal with this dataset. ( Peimin - master)  

# Analysis:
### Clean the data: 
1. isolate the month + year (Hanh - master branch)  
```{r load-data, cache = TRUE}
raw_data <- read.csv(here::here("data", "seek_australia_sample.csv"))
```

```{r clean-data}
# choose the needed columns only
data_cleaned <- raw_data %>%
  select(3:9)

# separate the post-date column to get the date out
data_cleaned <- data_cleaned %>%
  separate (post_date, into = c("post_date", "tz"), sep = "T")

# convert the post_date column from character to date format
data_cleaned %>%
  mutate_at(vars(post_date), as.Date, format = "%Y-%m-%d")
```

2. Missing values  (Hanh - Hanh-Ngo branch)  

### Analyze the change in demand by job category) (Peimin-Peimin branch)
```{r jobctegory}
library(kableExtra)
jobcategory <- data_cleaned  %>% dplyr::select(post_date,
                                               category) 
kable(jobcategory, caption = "2018 Job Category in Australia") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
head(jobcategory)
```


```{r category_months}
category_months <- jobcategory %>%  separate(post_date, c("Year","Month","day"))%>%
select(-"Year", -"day") 

```

```{r monthly_dirtribution}
category <- category_months  %>%
  count(Month,category,sort = TRUE, name = "number")%>%
  arrange(Month) %>%
   ggplot(aes(x = Month, y = number)) + 
   geom_col(position = "dodge") +
   facet_wrap(~category, ncol = 5)+ 
   theme_bw(base_size = 6.5)+
   theme(axis.text.x = element_text(angle = 50, vjust = 1)) + 
   ggtitle("Monthly job category distribution") 

category
```

```{r Month_number}
monthly <- category_months %>% 
  count(category,Month,sort = TRUE, name = "number") %>%
  arrange(Month) %>%
  group_by(Month) %>%
  top_n(5) %>%
   ggplot() +
  geom_point(mapping = aes(x = Month, y = number, color = category)) +
  
   theme_bw(base_size = 11.5)+
   theme(axis.text.x = element_text(angle = 50, vjust = 1)) 
   
monthly
```



### Analyze the change in demand by city + life quality  
1. Change in demand (Hanh)    
2. How does it reflect life quality? (Dewi)

### Analyze the change in demand by job type (Justin’s branch)

# Conclusion: (Justin - master branch)




